### 2023-01-27 definitions.tex
details

Properties in well-structured transition systems are often given as upward- or downward closed sets \cite{DBLP:journals/iandc/AbdullaCJT00, DBLP:journals/tcs/FinkelS01}.
%  Parosh Aziz Abdulla, Karlis Cerans, Bengt Jonsson & Yih-Kuen Tsay (1996): General Decidability Theorems for Infinite-State Systems. In: Proc. LICS 1996, IEEE Computer Society Press, pp. 313–321,
%  Alain Finkel & Philippe Schnoebelen (2001): Well-structured transition systems everywhere! Theor. Comput. Sci. 256(1-2), pp. 63–92, doi:10.1016/S0304-3975(00)00102-X.
Transfering the abstract resilience problems into this framework,
it is therefore reasonable to demand that both propositions, SAFE and BAD, are given by 
upward-closed or 
downward-closed sets.

We assume that the safety property is given by an upward-closed set and the bad condition by a decidable downward-closed set.

% \textcolor{red}{Seems like a reasonable assumption to me.}

From these considerations, we formulate instances of the abstract resilience problems for well-
structured transition systems.

@@

\problemx{$k$-resilience problem for WSTS}
{A state $s$ of a WSTS $(S,\rightarrow, \leq)$, an upward-closed set $U$ with a given basis, a decidable downward-closed set $D$.}
{$\forall s' ~ (s \rightarrow^* s') \wedge s' \in D  \implies \exists s'' \in U ~ s' \rightarrow^{\leq k} s''$ ?\newline}
%
plus simple : la $(D,U)$-resilience: $D \rightarrow^{\leq k} U$ ?  \alain{ça signifie pour tout $d \in D$ il existe un chemin de longueur $\leq k$ qui arrive dans un $u \in U$. c'est plus simple et plus général et ça ne demande moins de décider l'accessibilité....comment faire pour vérifier pour tous les (D,U) ? sinon ça dépend tjs de l'état intial $s$ et s'il faut vérifier pour tous les $s$....}

@@

\textcolor{blue}{
justification des definitions: si on enlève la condition  an upward-closed set $U$ (extended resilience $U$ qq) reachability reduces to resilience (avec $U=\{x\}$ donc indecidable pour les modèles à reachability indécidable.
}

>>>>

The $k$-resilience problem is to decide whether from a bad state, is it possible to find a path of length smaller than a given $k$ to reach a safe state. Let us define two sets BAD and SAFE. From all states $b \in BAD$ does there exist a path of length smaller than $k$ that reaches a state $s \in SAFE$ ?

It can be formally stated as follows :

\problemx{$k$-resilience problem}
{A transition system $(S,\rightarrow)$, two sets $SAFE, BAD \subseteq S$.}
{$BAD \longrightarrow^{\leq k} SAFE$ ?\newline}
%

For most of transition systems, this problem is decidable if both SAFE and BAD are finite. When one of the two sets is infinite, it generally becomes undecidable. So we restrict the framework to WSTS and sets SAFE and BAD will be upward or downward closed. In \cite{DBLP:conf/gg/Ozkan22} BAD is downward closed and SAFE is upward closed. But there are other possible cases: for instance, the well-known mutual exclusion property is often modelized in a $d$-counters machine by the property that a counter $c_{mutex}$ must be bounded by (usually) one. Then, the set $SAFE =  \{c_{mutex} \leq 1\} \times \mathbb{N}^{d-1}$ is downward closed and $BAD =\{c_{mutex} \geq 2\} \times  \mathbb{N}^{d-1} $ is the upward closed complementary of SAFE.
%

Most decidable properties in well-structured transition systems are given as upward or downward closed sets \cite{DBLP:journals/iandc/AbdullaCJT00, DBLP:journals/tcs/FinkelS01}.
%  Parosh Aziz Abdulla, Karlis Cerans, Bengt Jonsson & Yih-Kuen Tsay (1996): General Decidability Theorems for Infinite-State Systems. In: Proc. LICS 1996, IEEE Computer Society Press, pp. 313–321,
%  Alain Finkel & Philippe Schnoebelen (2001): Well-structured transition systems everywhere! Theor. Comput. Sci. 256(1-2), pp. 63–92, doi:10.1016/S0304-3975(00)00102-X.
Transfering the abstract resilience problems into this framework,
it is therefore reasonable to demand that both propositions, SAFE and BAD, are given by 
upward-closed or downward-closed sets.

\begin{tabular}{ l c r }
   BAD/SAFE & Up & Down \\
   Up & dec ? & dec ? \\
   Down & dec & ? \\
 \end{tabular}

We first assume that the safety property is given by an upward-closed set and the bad condition by a decidable downward-closed set. 
\alain{très discutable. pourquoi pas supposer des conditions plus gnérales ? perd-on la décidabilité ?}
% \textcolor{red}{Seems like a reasonable assumption to me.}

%From these considerations, we formulate instances of the abstract resilience problems for well-
%structured transition systems.

@@

\problemx{$k$-resilience problem for WSTS}
{A state $s$ of a WSTS $(S,\rightarrow, \leq)$, an upward-closed set $U$ with a given basis, a decidable downward-closed set $D$.}
{$D \longrightarrow^{\leq k} U$ ?\newline}
%
{$\forall s' ~ (s \rightarrow^* s') \wedge s' \in D  \implies \exists s'' \in U ~ s' \rightarrow^{\leq k} s''$ ?\newline}
%
 \alain{ça signifie pour tout $d \in D$ il existe un chemin de longueur $\leq k$ qui arrive dans un $u \in U$. c'est plus simple et plus général et ça ne demande moins de décider l'accessibilité....comment faire pour vérifier pour tous les (D,U) ? sinon ça dépend tjs de l'état intial $s$ et s'il faut vérifier pour tous les $s$....}

ou plus simple : $D \rightarrow^{\leq k} U$ ?

@@

%\textcolor{blue}{
%justification des definitions: si on enlève la condition  an upward-closed set $U$ (extended resilience $U$ qq) reachability reduces to resilience (avec $U=\{x\}$ donc indecidable pour les modèles à reachability indécidable.
%}
